---
title: "Data Quality"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
theme_set(theme_minimal())
options(scipen = 999)
```

# Data Quality Checks

Objectives:

-   Understand how many census tracts are present for all years of data

-   Understand how much data is missing by

    -   Census Tracts
    -   Year
    -   State

```{r}
original_df <- readRDS("outputs/combined.RDS")
dim(original_df)
df <- df[!duplicated(df), ]
dim(df)
colnames(df)
```

### Data Missing for PLACES Tracts:

```{r}
length(unique(df$GEOID))
length(unique(acs_2019$GEOID))
length(unique(redlining$GEOID))
length(unique(ltdb$GEOID))

ltdb %>%
  select(GEOID) %>%
  unique() %>%
  mutate(nchar = nchar(GEOID)) %>%
  group_by(nchar) %>%
  count()
```

```{r}
df %>%
  group_by(GEOID) %>%
  summarize(`Number of NAs in PLACES` = sum(is.na(ARTHRITIS)),
            `Number of Tracts` = n()) %>%
  ungroup() %>%
  group_by(`Number of NAs in PLACES`, `Number of Tracts`) %>%
  count() %>%
  mutate(Missing = ifelse(`Number of NAs in PLACES` == 0, "None", "All")) %>%
  ggplot() +
  geom_col(aes(x = as.factor(`Number of Tracts`), 
               y = `n`, 
               fill = Missing),
           position = "dodge") + 
  scale_y_log10() +
  labs(title = "Number of Tracts without PLACES data",
       y = "# Tracts (Log 10)", 
       x = "Tract Frequency") 
```

```{r}
df %>%
  group_by(GEOID) %>%
  summarize(`Number of NAs in PLACES` = sum(is.na(ARTHRITIS)),
            `Number of Tracts` = n()) 
  ungroup() %>%
  group_by(`Number of NAs in PLACES`, `Number of Tracts`) %>%
  count() %>%
  mutate(Missing = ifelse(`Number of NAs in PLACES` == 0, "None", "All"))  %>%
  ungroup() %>%
  group_by(Missing) %>%
  summarise(count = sum(n)) %>%
  mutate(total = length(unique(df$GEOID)),
         percent = count / total)
```

```{r}
df %>%
  group_by(GEOID) %>%
  summarize(`Number of NAs in PLACES` = sum(is.na(ARTHRITIS)),
            `Number of Tracts` = n()) %>%
  filter(`Number of NAs in PLACES` > 0)
  ungroup() %>%
  group_by(`Number of NAs in PLACES`, `Number of Tracts`) %>%
  count() %>%
  mutate(Missing = ifelse(`Number of NAs in PLACES` == 0, "None", "All")) %>%
  filter(Missing == "All")
```

**Slightly less than half the tracts are missing PLACES data.**

```{r}
missing_tracts <- df %>%
  group_by(GEOID) %>%
  summarize(`Number of NAs in PLACES` = sum(is.na(ARTHRITIS))) %>%
  filter(`Number of NAs in PLACES` > 0)

df %>%
  filter(GEOID %in% missing_tracts$GEOID) %>%
  group_by(STATE, YEAR) %>%
  count()
```

## Data Missing Across Years

```{r}
# Number of Tracts each decade
df %>%
  select(YEAR, GEOID) %>%
  group_by(YEAR) %>%
  count() %>%
  ggplot(aes(x = as.factor(YEAR), y = n, fill = as.factor(YEAR))) +
  geom_col(show.legend = F) +
  labs(x = "Year", y = "", title = "Number of Census Tracts by Year")
```

PLACES data missing across years

```{r}
df %>%
  select(YEAR, GEOID) %>%
  group_by(YEAR) %>% 
  mutate(missing = ifelse(GEOID %in% missing_tracts$GEOID, TRUE, FALSE)) %>%
  group_by(YEAR, missing) %>%
  count() %>%
  ggplot(aes(x = YEAR, y = n, fill = missing)) +
  geom_col() +
  labs(y = "# of CTs",
       x = "Year",
       fill = "PLACES data missing?") +
  theme(legend.position = "top")
```

## Redlining data intersection

```{r}
ltdb %>%
  select(GEOID, YEAR) %>%
  unique() %>%
  mutate(redlining = ifelse(GEOID %in% redlining$GEOID, TRUE, FALSE),
         places = ifelse(GEOID %in% places$TractFIPS, TRUE, FALSE)) %>%
  group_by(YEAR, redlining, places) %>%
  count() %>%
  ggplot(aes(x = YEAR, y = n, fill = redlining, alpha = places)) +
  geom_col(position = "dodge", 
           aes(color = places)) 


```

```{r}
need_to_clean <- df %>%
  select(YEAR, GEOID) %>%
  group_by(YEAR, GEOID) %>%
  count() %>% filter(n>1)

df %>% select(1:4) %>%
  filter(GEOID %in% need_to_clean$GEOID) %>%
  group_by(STATE, COUNTY, YEAR) %>% 
  count()
```

## Redlining Data - Duplicated Tracts in Multiple Cities

-   Temporarily corrected by averaging the HOLC grades for tracts with multiple values and removing NAs

```{r}
# Reading in Redlining Data
redlining <- read.csv("raw data/All_Population_Weighted_HOLC_Grades_pt2_Threshold.csv")

# Identifying tracts with GEOIDs recurring more than once
duplicates <- redlining %>%
  group_by(GEOID) %>%
  count() %>%
  filter(n > 1)

# Subsetting redlining data based on identified IDs
duplicates_all <- redlining %>%
  filter(GEOID %in% duplicates$GEOID)
```

### Duplicated Redlining Tracts Visualizations

```{r}
duplicates_all %>%
  group_by(city, state) %>%
  summarize(NAs = sum(is.na(holc_grade_pop)),
            `Total CTs` = n()) %>%
  pivot_longer(names_to = "Type", values_to = "Count", cols = c(NAs, `Total CTs`)) %>%
  
  ggplot(aes(x = reorder(city,Count), y = Count)) +
  geom_col(aes(fill = state, color = Type), 
           alpha = 0.4,
           position = "identity") +
  scale_color_manual(values = c("black", rgb(0,0,0,0)))+
  coord_flip() +
  labs(x = "# of CTs", y = "City", title = "Duplicated CTs")


duplicates_all %>%
  group_by(city, state) %>%
  summarize(NAs = sum(is.na(holc_grade_pop)),
            `Total CTs` = n()) %>%
  ungroup() %>%
  group_by(state) %>%
  summarize(NAs = sum(NAs),
            Total = sum(`Total CTs`))


duplicates_all %>%
  group_by(GEOID) %>%
  summarize(NAs = sum(is.na(holc_grade_pop)),
            `Total CTs` = n()) %>%
  ungroup() %>%
  group_by(NAs, `Total CTs`) %>%
  count() %>%
  ggplot(aes(x = `Total CTs`, y = n, group = NAs)) +
  geom_col(aes(fill = as.factor(NAs)), position = "dodge") +
  geom_text(aes(label = n), 
            position = position_dodge(width = 1), size = 3
            ) +
  labs(fill = "Number of NAs",
       x = "Number of Duplicated Tracts",
       y = "") +
  theme(axis.text.y = element_blank(),
        legend.position = "top")

```
