---
title: "Census-Based Gentrification Methods"
output: 
  html_notebook: 
    toc: true
    toc_depth: 3
    toc_float: true
    css:  "www/app.css"
runtime: shiny
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(tmap)
library(tigris)
library(sf)
library(MITREShiny)
library(DT)
library(plotly)
library(rgdal)

tracts <- readRDS("outputs/geo_tracts.RDS")
st_crs(tracts) = 4326

indicators <- readRDS("outputs/indicators.RDS") %>%
  mutate(STATE_FIP = substr(GEOID, 1,2),
         COUNTY_FIP = substr(GEOID, 3,5)) 
outcomes <- readRDS("outputs/outcomes.RDS")

holc_colors <- c(
  "A" = "#759e64", # green
  "B" = "#75a3b6", # blue
  "C" = "#c5c473", # yellow
  "D" = "#c27e8d" # red
)

gent_colors <- c(
  "#D0E1D4",  "#0C6291", "#C05746" 
)
fipsnames_cw <- unique(indicators[,c("STATE", "COUNTY", "STATE_FIP", "COUNTY_FIP")])

theme_set(theme_minimal()) +
  theme(plot.title = element_text(hjust=0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

# Introduction

The purpose of this document is to help users understand how the measures and indicators used in census-based gentrification affect which areas are classified as gentrified. In even as early as 1986, George Galster & Stephen Peacock reported "great sensitivity in the number and location of gentrified tracts to the definition chosen and stringency applied", when examining areal changes in Philadelphia between 1970 - 1980. [@galster1986] This definition explorer document hopes to visualize and elucidate some of those sensitivities.

Census-based gentrification definitions have often been used as a quantitative, quick way approximate the difficult concept of gentrification.

However, this has been little consensus among researchers in defining gentrification, much less in operationalizing the definition. In a systematic review of gentrification definitions used in epidemiologic research, 18 different indicators were used to capture and measure gentrification including Median Household Income, Education, Rent, Race, Home Value, Poverty, etc. [@Bhavsar2020]

Two common concepts are used in defining gentrification.

1.  Neighborhoods are **eligible** to be gentrified.
2.  Neighborhoods that meet threshold criteria are **determined** to be gentrified.

Indicators can be used to define eligibility, determination, or both.

In addition to areal indicators, time period of evaluation and reference geographic region are also important decisions which may impact the quantification of tracts as gentrified:

# Explore Indicators

```{r data_prep, echo=FALSE, message=FALSE}
# Additional Data prep to reduce processing later on
data <- indicators %>%
    select(GEOID, YEAR, STATE, COUNTY, `Total Pop`,
           `Median Rent`, 
           `Median Household Income`,
           `Percent with 4-year College Degree or More`, 
           `Median Home Value`,
           `Percent Above Poverty`,
           `Percent Non-Hispanic White`,
           `Percent Structures more than 30 years old`,
           `Households in neighborhood 10 years or less`) %>%
  gather(key = "Measure", value = "Value", -YEAR, -STATE, -COUNTY, -GEOID, -`Total Pop`) %>% 
  group_by(YEAR, STATE, COUNTY, Measure) %>%
  mutate(Rank = percent_rank(Value)) %>%
  ungroup()

# Generates scatter plot for all the measure functions used in the Explore Indicators Section
#'
#' @param data dataframe with columns `Rank Start`, `Rank End`, `Total Pop End` and `Label`
#' @return ggplot with defined aesthetics and geoms based on selected measure input `input$measure`

plot_measure_scatter <- function(data){
  plot <- data %>%
   ggplot(aes(x = `Rank Start`, y = `Rank End`)) +
      geom_point(aes(size = `Total Pop End`, color = Label), alpha = 0.7) +
      scale_color_manual(values = gent_colors) +
      theme(aspect.ratio = 1,
            legend.position = "bottom") +
    labs(title = paste(input$measure))
  
  return(plot)
}


#' Processes data to add categorical gentrification labels based on threshold inputs
#'
#' @param data dataframe with columns `Rank Start` and `Rank End`
#' @param input_eligibility User Defined input to determine the eligibility threshold
#' @param input_determination User Defined input to determine the determination threshold 
#' @return labelled datasframe with additional column `Label`

add_gentrification_label <- function(data, input_eligibility, input_determination){
  labelled <- data %>%
    mutate(Label = case_when(`Rank Start` > input_eligibility ~ "Not Eligible",
                             `Rank Start` <= input_eligibility & 
                               `Rank End` <= input_determination ~ "Still Susceptible",
                             `Rank Start` <= input_eligibility & 
                               `Rank End` > input_determination ~ "Positive Indicator"))
  return(labelled)
}
```

```{r core-explore, echo=FALSE, message=FALSE,}
# Input panel to select Year, State, County, and Measure
inputPanel(
    sliderInput("range", 
                label = "Select Year Range:",
                min = 1970, max = 2020, 
                value = c(2000, 2020), 
                step = 10, sep = ""),
    
    selectInput("state", 
                label = "Choose a State:",
                choices = sort(unique(data$STATE)), 
                selected = "Oregon"),
    
    renderUI({
      counties <- unique(data[data$STATE == input$state, "COUNTY"])
      selectInput("county", 
                  label = "Choose a County:",
                  choices = counties, 
                  selected = "Multnomah")
    }),
    
    selectInput("measure", 
                label = "Choose a Measure to Explore:",
                choices = unique(data$Measure),
                selected = "Median Household Income"
                )
  )


start <- reactive({
  data %>%
    filter(YEAR == input$range[1],
           STATE == input$state,
           COUNTY == input$county)
})

end <- reactive({
  data %>%
    filter(YEAR == input$range[2],
           STATE == input$state,
           COUNTY == input$county)
})

start_end <- reactive({
  inner_join(start(), end(),
               by = c("GEOID", "STATE", "COUNTY", "Measure"),
               suffix = c(" Start", " End")
    )
})

fips <- reactive({fipsnames_cw[fipsnames_cw$STATE == input$state & 
                             fipsnames_cw$COUNTY == input$county,]
    })
    

# Histogram + density plot to explore distribution of measures
splitLayout(
  
  renderPlot({
    data %>%
      filter(YEAR %in% input$range,
             STATE == input$state,
             COUNTY == input$county,
             Measure == input$measure) %>%
      ggplot(aes(x = Value)) +
      geom_histogram(aes(y = ..density..,
                         fill = as.factor(YEAR)),
                     alpha = 0.4, position = "dodge") + 
      geom_density(aes(color = as.factor(YEAR)), show.legend = FALSE) +
      labs(fill = "Year",
           y = "# of Tracts",
           title = paste(input$measure)) +
      theme(legend.position = "bottom",
            aspect.ratio = 1) 

  }),
  
  renderPlot({
    start_end() %>%
      filter(Measure == input$measure) %>%
      ggplot(aes(x = `Value Start`, y = `Value End`)) +
      geom_point(aes(size = `Total Pop End`),
                 alpha = 0.3) +
      geom_abline(slope = 1, intercept = 0, lty = "dotted", color = "gray") +
      theme(aspect.ratio = 1,
            legend.position = "bottom") +
      labs(title = paste(input$measure))
  })
)

```

# Define Indicators {.tabset}

Select the thresholds for each potential measure to define the eligibility and determination thresholds relative relative to the surrounding tracts in the county.

The selected thresholds will persist throughout the rest of the document.

-   Eligibility threshold characterizes which set of tracts can be considered candidates for gentrification.
-   Determination thresholds characterize what change is significant to capture neighborhood change.

## Median House Income

```{r mhi, echo=FALSE}
# Input panel information
inputPanel(
    sliderInput("eligibility_mhi", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_mhi", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )


# Generate MHI data set
mhi <- reactive({
  start_end() %>%
    filter(Measure == "Median Household Income") %>%
    add_gentrification_label(input_eligibility = input$eligibility_mhi, 
                             input_determination = input$determination_mhi)
})

tract_mhi <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,mhi(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    mhi() %>%
      plot_measure_scatter() 
  }),
  
   renderTmap({
        
        tm_shape(tract_mhi()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Median Rent

```{r rent, echo=FALSE}
inputPanel(
    sliderInput("eligibility_rent", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_rent", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )

# Generate MHI data set
rent <- reactive({
  start_end() %>%
    filter(Measure == "Median Rent") %>%
    add_gentrification_label(input_eligibility = input$eligibility_rent,
                             input_determination = input$determination_rent)
})

tract_rent <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,rent(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    rent() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        
        tm_shape(tract_rent()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Higher Education

```{r edu, echo=FALSE}
inputPanel(
    sliderInput("eligibility_edu", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_edu", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )

# Generate MHI data set
edu <- reactive({
  start_end() %>%
    filter(Measure == "Median Household Income") %>%
    add_gentrification_label(input_eligibility = input$eligibility_edu,
                             input_determination = input$determination_edu)
})

tract_edu <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,edu(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    edu() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        tm_shape(tract_edu()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Median Home Value

```{r mhv, echo=FALSE}
inputPanel(
    sliderInput("eligibility_mhv", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_mhv", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )


# Generate MHI data set
mhv <- reactive({
  start_end() %>%
    filter(Measure == "Median Home Value") %>%
    add_gentrification_label(input_eligibility = input$eligibility_mhv,
                             input_determination = input$determination_mhv)
})

tract_mhv <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,mhv(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    mhv() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        tm_shape(tract_mhv()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Percent Non-Hispanic White

```{r nhw, echo=FALSE}
inputPanel(
    sliderInput("eligibility_nhw", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_nhw", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )

# Generate NHW data set
nhw <- reactive({
  start_end() %>%
    filter(Measure == "Percent Non-Hispanic White") %>%
    add_gentrification_label(input_eligibility = input$eligibility_edu,
                             input_determination = input$determination_edu)
})

tract_edu <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,edu(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    edu() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        tm_shape(tract_edu()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Housing Structure Age

```{r hsa, echo=FALSE}
inputPanel(
    sliderInput("eligibility_hsa", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_hsa", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )

# Generate NHW data set
hsa <- reactive({
  start_end() %>%
    filter(Measure == "Percent Structures more than 30 years old") %>%
    add_gentrification_label(input_eligibility = input$eligibility_hsa,
                             input_determination = input$determination_hsa)
})

tract_hsa <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,hsa(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    hsa() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        tm_shape(tract_hsa()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Household Tenure

```{r ten, echo=FALSE}
inputPanel(
    sliderInput("eligibility_ten", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_ten", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )

# Generate tenure data set
ten <- reactive({
  start_end() %>%
    filter(Measure == "Households in neighborhood 10 years or less") %>%
    add_gentrification_label(input_eligibility = input$eligibility_ten,
                             input_determination = input$determination_ten)
})

tract_ten <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,ten(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    ten() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        tm_shape(tract_ten()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
)
```

## Percent Above FPL

```{r fpl, echo=FALSE}
inputPanel(
    sliderInput("eligibility_pov", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_pov", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )

# Generate tenure data set
pov <- reactive({
  start_end() %>%
    filter(Measure == "Percent Above Poverty") %>%
    add_gentrification_label(input_eligibility = input$eligibility_pov,
                             input_determination = input$determination_pov)
})

tract_pov <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,pov(), by = "GEOID")
      })

# Output
splitLayout(
  renderPlot({
    pov() %>%
      plot_measure_scatter()
  }),
  
   renderTmap({
        tm_shape(tract_pov()) +
          tm_polygons(col = "Label", 
                      palette = gent_colors,
                      alpha = 0.4)
      })
) 
```

# Select Indicators

Selection of indicators should be guided by local expertise in identifying the relevant indicators of gentrification within that community.

Based on the @Bhavsar2020 paper, the following were the most commonly used indicators across 24 papers examined in a recent literature review.

| Indicator Type                       | N = 24 (%) |
|:-------------------------------------|:-----------|
| Median Household income              | 17 (70.8%) |
| Education                            | 17 (70.8%) |
| Rent                                 | 12 (50%)   |
| Home Value                           | 6 (25%)    |
| Race                                 | 6 (25%)    |
| Poverty                              | 4 (16.7%)  |
| Housing Stock Age                    | 3 (12.5%)  |
| Composite Measure (SEIFA, IMD, NCGS) | 3 (12.5%)  |
| Immigrant Population %               | 2 (8.4%)   |
| Occupation                           | 2 (8.4%)   |
| Residential Mobility                 | 2 (8.4%)   |

Some additional measures used (though only appearing once among the reviewed literature) included Non-family households, Owner Occupied, Unemployment, Dollar Amount for Improvement Loans, and Mean Dollar Amount of Home Loans.

Our tool defaults to the top 3 most frequently used indicators of gentrification, but includes the top 7 most commonly used indicators.

```{r, echo = FALSE}


  checkboxGroupInput("measures",
                     label = "Select which measures you would like to include in your gentrification definition:",
                     choices = unique(data$Measure),
                     selected = unique(data$Measure)[1:3])



add_gentrification_label_full <- function(data, measure, input_eligibility, input_determination){
  data %>%
    mutate(Label = case_when(Measure == measure & `Rank Start` > input_eligibility ~ "Not Eligible",
                             Measure == measure & `Rank Start` <= input_eligibility & 
                               `Rank End` <= input_determination ~ "Still Susceptible",
                             Measure == measure & `Rank Start` <= input_eligibility & 
                               `Rank End` > input_determination ~ "Positive Indicator",
                             Measure != measure ~ Label))
}

start_end_labelled <- reactive({
  start_end() %>%
    mutate(Label = "")%>%
    add_gentrification_label_full("Median Rent", 
                                  input$eligibility_rent, 
                                  input$determination_rent) %>%
    add_gentrification_label_full("Median Household Income", 
                                  input$eligibility_mhi, 
                                  input$determination_mhi ) %>%
    add_gentrification_label_full("Percent with 4-year College Degree or More",
                                  input$eligibility_edu, 
                                  input$determination_edu) %>%
    add_gentrification_label_full("Median Home Value",
                                  input$eligibility_mhv, 
                                  input$determination_mhv) %>%
    add_gentrification_label_full("Percent Above Poverty",
                                  input$eligibility_pov, 
                                  input$determination_pov) %>%
    add_gentrification_label_full("Percent Non-Hispanic White",
                                  input$eligibility_nhw, 
                                  input$determination_nhw) %>%
    add_gentrification_label_full("Percent Structures more than 30 years old",
                                  input$eligibility_hsa, 
                                  input$determination_hsa) %>%
    add_gentrification_label_full("Households in neighborhood 10 years or less",
                                  input$eligibility_ten, 
                                  input$determination_ten) %>%
    filter(Measure %in% input$measures) %>%
    select(GEOID, STATE, COUNTY, Measure, Label) %>%
    spread(key = "Measure", value = "Label")
})

measures <- reactive({
  start_end_labelled() %>%
    mutate(`Positive Indicator` = as.integer(rowSums(. == "Positive Indicator")),
           `Not Eligible` = as.integer(rowSums(. == "Not Eligible")),
           `Still Susceptible` = as.integer(rowSums(. == "Still Susceptible")))
})


tract_measures <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,measures(), by = "GEOID")
      })

renderTmap({
  tm_shape(tract_measures()) +
    tm_polygons(col = "Still Susceptible",
                alpha = 0.5,
                palette = c("#FFFFFF", gent_colors[3]),
                legend.format = list(digits = 0)
                ) +
    tm_shape(tract_measures()) +
    tm_polygons(col = "Positive Indicator",
                alpha = 0.5,
                palette = c("#FFFFFF", gent_colors[2]),
                legend.format = list(digits = 0),
                popup.vars = c("Positive Indicator",
                               "Still Susceptible",
                               "Not Eligible"),
                popup.format = list(digits = 0))
      }) 

renderDataTable(
  datatable({
    measures()
  },
  extensions = 'Buttons',
     options = list(
       dom = 'lBtip',
       buttons = c('copy', 'excel', 'csv'),
       pageLength = 10,
       lengthMenu = list(c(10,25,50,-1),
                         c('10', '25', '50', 'All'))
     ),
     class = "display"
  )
)
# 


```

# Compare Results

Given the selected measures and thresholds to mark gentrification, this tool allows for comparisons with selected health outcomes.

The health outcomes data come from the [U. S. Small-area Life Expectancy Estimates Project (USALEEP)](https://www.cdc.gov/nchs/nvss/usaleep/usaleep.html), and from CDC's [PLACES: Local Data for Better Health](https://chronicdata.cdc.gov/browse?category=500+Cities+%26+Places&sortBy=newest&utf8).

To use this section, select a health outcome of interest from the drop-down list. The following visuals will display a map colored by the number and type of gentrification indicators, with bubbles representing the prevalence of the the selected health outcome.

The violin plots below show the relationship between the prevalence distribution and the number of Positive Gentrification Indicators (left) and the number of Still Eligible Gentrification Indicator (right).

```{r, echo = FALSE}
outcomes <- outcomes %>%
  select(GEOID, `Life Expectancy`,
         BINGE, CSMOKING, BPHIGH, CANCER, CASTHMA, DIABETES, HIGHCHOL, MHLTH, PHLTH, LPA, SLEEP, ACCESS2, DENTAL) %>%
  unique()

colnames(outcomes) <- c("GEOID",
                        "Life Expectancy",
                        "Binge Drinking",
                        "Current Smoker",
                        "High Blood Pressure",
                        "Cancer",
                        "Asthma",
                        "Diabetes",
                        "High Cholesterol",
                        "Poor Mental Health",
                        "Poor Physical Health",
                        "Limited Physical Activity",
                        "Insufficient Sleep",
                        "Uninsured",
                        "Dental Visits")

outcomes <- outcomes %>%
  gather(key = "Health Outcomes",
         value = "Prevalence",
         -GEOID, na.rm = TRUE)


inputPanel(
    selectInput("brfss", 
                label = "Choose a Health Outcome to Explore:",
                choices = unique(outcomes$`Health Outcomes`),
                selected = "Binge Drinking"
                )
  )

outcomes_measures <- reactive({
  measures() %>%
    left_join(outcomes, by = "GEOID") %>%
    filter(`Health Outcomes` == input$brfss)
})

tracts_outcomes_measures <- reactive({
  tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,outcomes_measures(), by = "GEOID")

})

renderTmap({

  tm_shape(tract_measures()) +
    tm_polygons(col = "Still Susceptible",
                alpha = 0.5,
                palette = c("#FFFFFF", gent_colors[3]),
                legend.format = list(digits = 0)
                ) +
    tm_shape(tract_measures()) +
    tm_polygons(col = "Positive Indicator",
                alpha = 0.5,
                palette = c("#FFFFFF", gent_colors[2]),
                legend.format = list(digits = 0),
                popup.vars = c("Positive Indicator",
                               "Still Susceptible",
                               "Not Eligible"),
                popup.format = list(digits = 0)) +
    tm_shape(tracts_outcomes_measures()) +
    tm_bubbles(col = "Prevalence",
               size = "Prevalence",
               palette = "Greens",
               alpha = 0.5,
               scale = 0.5)
})

splitLayout(
  renderPlot({
    
    counts <- outcomes_measures() %>%
      group_by(`Positive Indicator`) %>%
      summarize(n = n(),
                Prevalence = mean(Prevalence, na.rm = T))
    
    outcomes_measures() %>%
      ggplot(aes(y = Prevalence, 
                 x = factor(`Positive Indicator`))) +
      geom_violin(aes(fill = factor(`Positive Indicator`)),
                  alpha = 0.6, 
                  color = "white",
                  show.legend = FALSE) +
      geom_text(data = counts, aes(label = n)) +
      labs(x = "Number of Positive Indicators of Gentrification",
           y = paste("Prevalence"),
           title = paste(input$brfss, "by # of Gentrification Indicators"))
  }),
  
  renderPlot({
    
    counts <- outcomes_measures() %>%
      group_by(`Still Susceptible`) %>%
      summarize(n = n(),
                Prevalence = mean(Prevalence, na.rm = T))
    
    outcomes_measures() %>%
      ggplot(aes(y = Prevalence,
                 x = factor(`Still Susceptible`))) +
      geom_violin(aes(fill = factor(`Still Susceptible`)),
                  alpha = 0.6, 
                  color = "white",
                  show.legend = FALSE) +
      geom_text(data = counts, aes(label = n)) +
      labs(x = "Number of Still Susceptible Indicators of Gentrification",
           y = paste("Prevalence"),
           title = paste(input$brfss, "by # of Susceptibility Indicators"))
  })
)

```
