---
title: "Census-Based Gentrification Methods"
output: 
  html_notebook: 
    toc: true
    toc_depth: 3
    toc_float: true
    css:  "www/app.css"
runtime: shiny
bibliography: references.bib
---

```{r setup, include=FALSE}

library(ggplot2)
library(dplyr)
library(tidyr)
library(tmap)
library(tigris)
library(sf)
library(MITREShiny)
library(DT)

setwd("~/Documents/MIP/gentrification")
tracts <- readRDS("outputs/geo_tracts.RDS")

combined <- readRDS("outputs/combined.RDS") %>%
  mutate(YEAR = ifelse(YEAR == 2019, 2020, YEAR),
         STATE_FIP = substr(GEOID, 1,2),
         COUNTY_FIP = substr(GEOID, 3,5))

holc_colors <- c(
  "A" = "#759e64", # green
  "B" = "#75a3b6", # blue
  "C" = "#c5c473", # yellow
  "D" = "#c27e8d" # red
)

fipsnames_cw <- unique(combined[,c("STATE", "COUNTY", "STATE_FIP", "COUNTY_FIP")])

# holc <- read.csv("../raw data/All_Population_Weighted_HOLC_Grades_pt2_Threshold.csv",
#                  colClasses = c("character", "character", "character", "double"))
# lexp <- read.csv("../raw data/US_A.CSV",
#                  colClasses = c("character","character","character","character","double", "double","integer"),
#                  check.names = F)
# colnames(lexp) <- c("GEOID", "STATE", "COUNTY", "TRACT", "Life Expectancy", "SE", "Flag")

theme_set(theme_minimal())
```

## Introduction

The purpose of this document is to help users understand how the measures and indicators used in census-based gentrification affect which areas are classified as gentrified. In even as early as 1986, George Galster & Stephen Peacock reported "great sensitivity in the number and location of gentrified tracts to the definition chosen and stringency applied", when examining areal changes in Philadelphia between 1970 - 1980. [@galster1986] This definition explorer document hopes to visualize and elucidate some of those sensitivities.

Census-based gentrification definitions have often been used as a quantitative, quick way approximate the difficult concept of gentrification.

However, this has been little consensus among researchers in defining gentrification, much less in operationalizing the definition. In a systematic review of gentrification definitions used in epidemiologic research, 18 different indicators were used to capture and measure gentrification including Median Household Income, Education, Rent, Race, Home Value, Poverty, etc. [@Bhavsar2020]

Two common concepts are used in defining gentrification.

1.  Neighborhoods are **eligible** to be gentrified.
2.  Neighborhoods that meet threshold criteria are **determined** to be gentrified.

Indicators can be used to define eligibility, determination, or both.

In addition to areal indicators, time period of evaluation and reference geographic region are also important decisions which may impact the quantification of tracts as gentrified.

## Explore Indicators

The three most commonly used indicators @Bhavsar2020 used in quantitative definitions of gentrification are:

-   Median Household Income

-   Median Rent

-   Education (% with Bachelors Degree)


```{r core-explore, echo=FALSE}


  inputPanel(
    sliderInput("range", label = "Select Year Range:",
                min = 1970, max = 2020, value = c(2000, 2020), step = 10, sep = ""),
    
    selectInput("state", label = "Choose a State:",
                choices = sort(unique(combined$STATE)), selected = "Oregon"),
    
    renderUI({
      counties <- sort(unique(combined[combined$STATE == input$state, "COUNTY"]))
      selectInput("county", label = "Choose a County:",
                  choices = counties, selected = "Multnomah")
    })
  )
  
renderPlot({
  
  combined %>%
    select(YEAR, STATE, COUNTY, 
           `Median Rent`, 
           `Median Household Income, Total`,
           `Percent with 4-year College Degree or More`) %>%
    gather(key = "Feature", value = "Value", -YEAR, -STATE, -COUNTY) %>%
    filter(YEAR %in% input$range,
           STATE == input$state,
           COUNTY == input$county) %>%
    ggplot(aes(x = Value)) +
    geom_histogram(aes(fill = as.factor(YEAR)),
                   alpha = 0.6, position = "dodge") +
    facet_wrap(~ Feature, nrow = 1, ncol = 4, scales="free") +
    labs(fill = "Year",
         y = "# of Tracts") +
    theme(legend.position = "bottom",
          aspect.ratio = 0.8)


})

```

Eligibility threshold characterizes which set of tracts can be considered candidates for gentrification.

Determination thresholds characterize what change is significant to capture neighborhood change.

### Median Household Income

```{r mhi, echo=FALSE}

  inputPanel(
    sliderInput("eligibility_mhi", "Select Elgibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_mhi", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_mhi <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(`Median Household Income, Total`))
    })
    
    end_data_mhi <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(`Median Household Income, Total`))
    })
    
    labelled_data_mhi <- reactive({
      start_data_mhi() %>%
        inner_join(end_data_mhi(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(`Eligible (MHI)` = ifelse(Rank_start < input$eligibility_mhi, "Yes", "No"),
               `Determination (MHI)` = ifelse(Rank_end > input$determination_mhi & `Eligible (MHI)` == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (MHI)` = ifelse(`Eligible (MHI)` == "Yes" & `Determination (MHI)` == "Yes", 
                                   "Yes", "No"))
    })
  

    
    fips <- reactive({fipsnames_cw[fipsnames_cw$STATE == input$state & 
                             fipsnames_cw$COUNTY == input$county,]
    })
    
    tract_subset_mhi <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_mhi(), by = "GEOID")
      })
    
    splitLayout(
      renderPlot({
        labelled_data_mhi() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = `Eligible (MHI)`, alpha = `Determination (MHI)`, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        
        tm_shape(tract_subset_mhi()) +
          tm_polygons(col = "Gentrification Indicator (MHI)",
                      alpha = 0.6)
      })
      
    )
    
```

### Median Rent 

Including a Median Rent Indicator means that only tracts that started with a median rental value below some X threshold were eligible to gentrify, and if among those tracts, ended up above the top percentile

```{r rent, echo=FALSE}
  inputPanel(
    sliderInput("eligibility_rent", "Select Eligibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_rent", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_rent <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(`Median Rent`))
    })
    
    end_data_rent <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(`Median Rent`))
    })
    
    labelled_data_rent <- reactive({
      start_data_rent() %>%
        inner_join(end_data_rent(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(Eligible = ifelse(Rank_start < input$eligibility_rent, "Yes", "No"),
               Determination = ifelse(Rank_end > input$determination_rent & Eligible == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (Rent)` = ifelse(Eligible == "Yes" & Determination == "Yes", 
                                   "Yes", "No"))
    })
  
    tract_subset_rent <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_rent(), by = "GEOID")
    })
    
    splitLayout(
      renderPlot({ 
        labelled_data_rent() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = Eligible, alpha = Determination, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        
        tm_shape(tract_subset_rent()) +
          tm_polygons(col = "Gentrification Indicator (Rent)",
                      alpha = 0.4)
      })
    )
```

### Education 

Including an Education Indicator means that only tracts with a population above a pre-defined percentage threshold of residents with a 4-year College Degree or More can be considered gentrified.

```{r edu, echo=FALSE}
  inputPanel(
    sliderInput("eligibility_edu", "Select Eligibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1),
    
    sliderInput("determination_edu", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_edu <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(`Percent with 4-year College Degree or More`))
    })
    
    end_data_edu <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(`Percent with 4-year College Degree or More`))
    })
    
    labelled_data_edu <- reactive({
      start_data_edu() %>%
        inner_join(end_data_edu(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(Eligible = ifelse(Rank_start < input$eligibility_edu, "Yes", "No"),
              Determination = ifelse(Rank_end > input$determination_edu & Eligible == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (Education)` = ifelse(Eligible == "Yes" & Determination == "Yes", 
                                   "Yes", "No"))
    })
  
    tract_subset_edu <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_edu(), by = "GEOID")
    })
    
    splitLayout(
      renderPlot({ 
        labelled_data_edu() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = Eligible, alpha = Determination, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        tm_shape(tract_subset_edu()) +
          tm_polygons(col = "Gentrification Indicator (Education)",
                      alpha = 0.4)
      })
    )
```

### Additional Indicators

#### Median Home Value 

```{r mhv, echo=FALSE, collapse=TRUE}
  inputPanel(
    sliderInput("eligibility_mhv", "Select Eligibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_mhv", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_mhv <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(`Median Home Value`))
    })
    
    end_data_mhv <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(`Median Home Value`))
    })
    
    labelled_data_mhv <- reactive({
      start_data_mhv() %>%
        inner_join(end_data_mhv(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(Eligible = ifelse(Rank_start < input$eligibility_mhv, "Yes", "No"),
               Determination = ifelse(Rank_end > input$determination_mhv & Eligible == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (MHV)` = ifelse(Eligible == "Yes" & Determination == "Yes", 
                                   "Yes", "No"))
    })
  
    tract_subset_mhv <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_mhv(), by = "GEOID")
    })
    
    splitLayout(
      renderPlot({ 
        labelled_data_mhv() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = Eligible, alpha = Determination, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        
        tm_shape(tract_subset_mhv()) +
          tm_polygons(col = "Gentrification Indicator (MHV)",
                      alpha = 0.4)
      })
    )
```

#### Percent in Poverty

```{r, echo=FALSE, collapse=TRUE}
  inputPanel(
    sliderInput("eligibility_pov", "Select Eligibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_pov", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_pov <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(-`Percent in Poverty, Total`))
    })
    
    end_data_pov <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(-`Percent in Poverty, Total`))
    })
    
    labelled_data_pov <- reactive({
      start_data_pov() %>%
        inner_join(end_data_pov(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(Eligible = ifelse(Rank_start < input$eligibility_pov, "Yes", "No"),
               Determination = ifelse(Rank_end > input$determination_pov & Eligible == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (Poverty)` = ifelse(Eligible == "Yes" & Determination == "Yes", 
                                   "Yes", "No"))
    })
  
    tract_subset_pov <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_pov(), by = "GEOID")
    })
    
    splitLayout(
      renderPlot({ 
        labelled_data_pov() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = Eligible, alpha = Determination, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        
        tm_shape(tract_subset_pov()) +
          tm_polygons(col = "Gentrification Indicator (Poverty)",
                      alpha = 0.4)
      })
    )
```

#### Percent Non-Hispanic White

Find literature references for this...
eligible = areas that start out with relatively low white pop
gentrified = areas that end up with higher white pop

```{r, echo=FALSE, collapse=TRUE}
  inputPanel(
    sliderInput("eligibility_nhw", "Select Eligibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_nhw", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_nhw <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(`Percent Non-Hispanic White`))
    })
    
    end_data_nhw <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(`Percent Non-Hispanic White`))
    })
    
    labelled_data_nhw <- reactive({
      start_data_nhw() %>%
        inner_join(end_data_nhw(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(Eligible = ifelse(Rank_start < input$eligibility_pov, "Yes", "No"),
               Determination = ifelse(Rank_end > input$determination_pov & Eligible == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (Non-Hispanic White)` = ifelse(Eligible == "Yes" & Determination == "Yes", 
                                   "Yes", "No"))
    })
  
    tract_subset_nhw <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_nhw(), by = "GEOID")
    })
    
    splitLayout(
      renderPlot({ 
        labelled_data_nhw() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = Eligible, alpha = Determination, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        
        tm_shape(tract_subset_nhw()) +
          tm_polygons(col = "Gentrification Indicator (Non-Hispanic White)",
                      alpha = 0.4)
      })
    )
```

#### Housing Structure Age

The housing structure age variable is defined as the % of structures in a neighborhood that are more than 30 years old.
An area that starts out with a large percentage of old housing structures greater than 30 years may indicate areas that are eligible to gentrify.


```{r, echo=FALSE, collapse=TRUE}
  inputPanel(
    sliderInput("eligibility_hsa", "Select Eligibility Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1), 
    
    sliderInput("determination_hsa", "Select Determination Threshold",
                min = 0, max = 1, value = 0.5, step = 0.1)
  )
  
  
    start_data_hsa <- reactive({ 
      combined %>%
        filter(YEAR == input$range[1],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_start = percent_rank(`Percent Structures more than 30 years old`))
    })
    
    end_data_hsa <- reactive({  
      combined %>%
        filter(YEAR == input$range[2],
               STATE == input$state,
               COUNTY == input$county) %>%
        mutate(Rank_end = percent_rank(`Percent Structures more than 30 years old`))
    })
    
    labelled_data_hsa <- reactive({
      start_data_hsa() %>%
        inner_join(end_data_hsa(), by = "GEOID") %>% 
        select(Rank_start, Rank_end, GEOID, `Total Pop.y`) %>%
        mutate(Eligible = ifelse(Rank_start > input$eligibility_hsa, "Yes", "No"),
               Determination = ifelse(Rank_end < input$determination_hsa & Eligible == "Yes",
                                      "Yes", "No"),
               `Gentrification Indicator (Housing Structure Age)` = ifelse(Eligible == "Yes" & Determination == "Yes", 
                                   "Yes", "No"))
    })
  
    tract_subset_hsa <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(.,labelled_data_hsa(), by = "GEOID")
    })
    
    splitLayout(
      renderPlot({ 
        labelled_data_hsa() %>%
          ggplot(aes(x = Rank_start, y = Rank_end)) +
          geom_point(aes(color = Eligible, alpha = Determination, size = `Total Pop.y`)) +
          scale_alpha_manual(values = c(0.3, 1))+ 
          theme(aspect.ratio = 1) +
          labs(x = "Tract Percentile Rank in Start Decade",
               y = "Tract Percentile Rank in End Decade",
               size = "Population in End Decade")
        
      }),
      
      renderTmap({
        
        tm_shape(tract_subset_hsa()) +
          tm_polygons(col = "Gentrification Indicator (Housing Structure Age)",
                      alpha = 0.4)
      })
    )
```

## Output and Download

Census Tracts that meet the determination thresholds in Median Household Income, Median Rent, and Education are consequently classified as gentrified.

```{r, echo=FALSE}

inputPanel(
  checkboxInput("mhi", "Include Median Household Income Indicator", value = TRUE),
  
  checkboxInput("rent", "Include Median Rent Indicator", value = TRUE),
  
  checkboxInput("edu", "Include Education Indicator", value = TRUE),
  
  checkboxInput("mhv", "Include Median Home Value Indicator", value = FALSE),

  checkboxInput("pov", "Include Poverty Indicator", value = FALSE),
  
  checkboxInput("nhw", "Include Non-Hispanic White Indicator", value = FALSE),
  
  checkboxInput("hsa", "Include Housing Structure Age", value = FALSE)
)


  
  starter <- tibble(GEOID = "")
  
  right_join_if_true <- function(left_data, right_data, by, suffix, bool){
    if (bool) {
      right_join(left_data, right_data, by = by, suffix = suffix)
    } else {left_data}
  }
  
  combined_set <- reactive({
    starter %>%
      right_join_if_true(labelled_data_mhi(), by = "GEOID", suffix = c("", " MHI"), input$mhi) %>%
      right_join_if_true(labelled_data_rent(), by = "GEOID", suffix = c("", " Rent"), input$rent) %>%
      right_join_if_true(labelled_data_edu(), by = "GEOID", suffix = c("", " Edu"), input$edu) %>%
      right_join_if_true(labelled_data_mhv(), by = "GEOID", suffix = c("", " MHV"), input$mhv) %>% 
      right_join_if_true(labelled_data_pov(), by = "GEOID", suffix = c("", " Pov"), input$pov) %>%
      right_join_if_true(labelled_data_nhw(), by = "GEOID", suffix = c("", " NHW"), input$nhw) %>%
      right_join_if_true(labelled_data_hsa(), by = "GEOID", suffix = c("", " HSA"), input$hsa)

  })
    
    
  indicators <- reactive({
    combined_set() %>%
      select(GEOID, starts_with("Gentrification Indicator")) %>%
      mutate(`Number of Indicators` = rowSums(.[,-1] == "Yes"))
  })
  
  
  eligibility_ind <- reactive({
    combined_set() %>%
      select(GEOID, starts_with("Eligible")) %>%
      mutate(`Eligibility Indicators` = rowSums(.[,-1] == "Yes"))
  })                         

    
 combined_tract_subset <- reactive({tracts %>%
        filter(STATEFP == fips()$STATE_FIP,
               COUNTYFP == fips()$COUNTY_FIP) %>%
        merge(., indicators(), by = "GEOID") %>%
        merge(., eligibility_ind(), by = "GEOID")
    })

 
 renderTmap({
   tm_shape(combined_tract_subset()) +
     tm_polygons(col = "Number of Indicators",
                 alpha = 0.4) 
 })
 renderTmap({
   tm_shape(combined_tract_subset()) +
     tm_polygons(col = "Eligibility Indicators",
                 alpha = 0.4) 
 })
 renderDataTable(
   datatable(
     {indicators()},
     extensions = 'Buttons',
     options = list(
       dom = 'lBtip',
       buttons = c('copy', 'excel', 'csv'),
       pageLength = 10,
       lengthMenu = list(c(10,25,50,-1),
                         c('10', '25', '50', 'All'))
     ),
     class = "display"
   )
 )
```

## Gentrification in Context

Gentrification does not occur in an isolated context. Neil Smith's "rent gap" theory suggests neighborhoods in central areas that faced disinvestment may more subject to gentrification. In particular, his theory suggests that the gap between potential and actual rent prices drives gentrification. @smith1987

To place both modern and historic gentrification in context, we can examine present day patterns of gentrification alongside historic redlining practices.

+--------------------------+-------------------+-------------------------+
| HOLC Grade               | Numeric Code      | Original Map Color Code |
+==========================+===================+=========================+
| A (Best)                 | 1                 | Green                   |
+--------------------------+-------------------+-------------------------+
| B (Still Desirable)      | 2                 | Blue                    |
+--------------------------+-------------------+-------------------------+
| C (Definitely Declining) | 3                 | Yellow                  |
+--------------------------+-------------------+-------------------------+
| D (Hazardous)            | 4                 | Red                     |
+--------------------------+-------------------+-------------------------+

### Redlining

Here we have side by side maps examining gentrification indicators and historic redlining maps.

```{r, echo=FALSE}
indicators_holc <- reactive({combined_tract_subset() %>%
    merge(., end_data_mhi(), by = "GEOID")
})

splitLayout(
  renderTmap({
   tm_shape(indicators_holc()) +
     tm_polygons(col = "Number of Indicators",
                 alpha = 0.4)
  }),
  
  renderTmap({
    tm_shape(indicators_holc()) +
     tm_polygons(col = "holc_grade_pop",
                 palette = holc_colors,
                lwd = 1.2,
                alpha = 0.6, 
                title = "HOLC Grades")
  })
)


```

Combining the two maps above, we can look at the relationship between historic HOLC grades and the number of Gentrification indicators. Each circle represents a census tract, and the size represents the total population at the ending decade.

```{r, echo=FALSE}



renderPlot(
  indicators_holc() %>%
    ggplot() +
    geom_jitter(aes(x = holc_grade_pop, 
                   y = `Number of Indicators`, 
                   color = holc_grade_pop,
                   size = `Total Pop`), 
               alpha = 0.8,
               height = 0.1, 
               show.legend = FALSE) +
    scale_color_gradientn("HOLC Grade",
                           colors = holc_colors,
                           labels = c("A","B","C","D"),
                           breaks = c(1:4),
                           limits = c(1,4),
                           na.value = "transparent") +
    labs(x = "HOLC Grade (Redlining)",
         y = "Number of Gentrification Indicators")
)



renderPlot(
  indicators_holc() %>%
    ggplot() +
    geom_jitter(aes(x = holc_grade_pop, 
                   y = `Eligibility Indicators`, 
                   color = holc_grade_pop,
                   size = `Total Pop`), 
               alpha = 0.8,
               height = 0.1, 
               show.legend = FALSE) +
    scale_color_gradientn("HOLC Grade",
                           colors = holc_colors,
                           labels = c("A","B","C","D"),
                           breaks = c(1:4),
                           limits = c(1,4),
                           na.value = "transparent") +
    labs(x = "HOLC Grade (Redlining)",
         y = "Number of Gentrification Indicators")
)




# 
# renderPlot(
#   indicators_holc() %>%
#     ggplot() +
#     geom_point(aes(x = holc_grade_pop, 
#                    y = `MHLTH`, 
#                    color = factor(`Number of Indicators`),
#                    size = `Total Pop`),
#                alpha = 0.4) +
#     labs(x = "HOLC Grade (Redlining)",
#          y = "Number of Gentrification Indicators",
#          y = "Poor Mental Health Prevalence")
# )

```
