---
title: "ACS Flows"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("mapdeck")
library(tidyverse)
library(tidycensus)
library(mapdeck)
```

## ACS Migration Flows

Can we use migration flow data to better understand displacement?

<https://www.census.gov/data/developers/data-sets/acs-migration-flows.html>

**Definitions:**

-   `Moved In`: Reference -\> Second County
-   `Moved Out`: Reference \<- Second County
-   `Moved Net`: Moved In - Moved Out

```{r}
wch_flows <- get_flows(
  geography = "county",
  state = "OR",
  county = "Multnomah",
  year = 2018,
  geometry = T
  )



```

```{r}
top_move_in <- wch_flows %>% 
  filter(!is.na(GEOID2), variable == "MOVEDOUT") %>% 
  slice_max(n = 25, order_by = estimate) %>% 
  mutate(
    width = estimate / 500,
    tooltip = paste0(
      scales::comma(estimate * 5, 1),
      " people moved from ", str_remove(FULL2_NAME, "Metro Area"),
      " to ", str_remove(FULL1_NAME, "Metro Area"), " between 2014 and 2018"
      )
    )

top_move_in %>% 
  mapdeck(style = mapdeck_style("dark"), pitch = 45) %>% 
  add_arc(
    origin = "centroid1",
    destination = "centroid2",
    stroke_width = "width",
    auto_highlight = TRUE,
    highlight_colour = "#8c43facc",
    tooltip = "tooltip"
  )
```
