---
title: "Places Analysis"
author: "AJ Liberatore"
date: "4/12/2022"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(tigris)
library(sp)
library(sf)
library(ggplot2)
library(dplyr)
library(shiny)
library(leaflet)
library(gridExtra)
library(EnvStats)
theme_set(theme_minimal())
options(scipen = 999)
```

# Analysis of Places Geography for HOLC Graded Cities

Objectives:

- Find out how many places intersect with the original HOLC boundaries. This will give us an idea of the places that encapsulate cities.

- How much bigger are the combined places than the original HOLC boundaries? We expect that populations have expanded and/or migrated from the original boundaries, and expect the combined places geographies to reflect that.

- Add tract lines within the places for plots (lots of plots). Thus, every place/combined place geography is broken down by census tract. Additional granularity is always a good move.

- See if setting cb to TRUE when pulling places makes changes. This will pull a more detailed shapefile than the default (cb = FALSE). Ideally, we will get a better result, but generalizing should do for now.

## Loading necessary data
```{r, echo = FALSE, results='hide'}
# Lookup table for states with GEOID and 2 letter abbreviation
state_lookup <- tigris::states() %>%
  select(GEOID, STUSPS) %>%
  st_drop_geometry()

# Loads places data with geography
places <- readRDS("../outputs/geo_places.RDS")

# Loads tracts data with geography
tracts <- readRDS("../outputs/geo_tracts.RDS")

# Loads holc data with tracts, city, state, and their holc grade
holc <- read.csv("../raw data/All_Population_Weighted_HOLC_Grades_pt2_Threshold.csv", 
                 colClasses = c("GEOID" = "character"))

# Loads tracts data with geography and the place which it is in
places_tracts <- readRDS("../outputs/final_places_tracts.RDS") %>%
  left_join(state_lookup, by = c("STATEFP" = "GEOID"))

# Loads holc data with tracts, city, place, and their holc grade
holc_geo <- readRDS("../outputs/holc_geo.RDS")

# Loads city data with combined places geography
holc_place <- readRDS("../outputs/holc_place.RDS")

# Loads place data with geography for places with population estimate greater than 50k in 2019
places_pop_gt_50k <- readRDS("../outputs/places_pop_gt_50k.RDS")

# Loads population estimates for tracts
all_acs_pops <- readRDS("../outputs/all_acs_pops.RDS")
```

```{r, echo=FALSE, results='hide'}
# Calculates area of combined places data for cities, m^2
place_areas <- holc_place %>%
  dplyr::mutate(place_area = as.numeric(sf::st_area(geometry))) %>%
  st_drop_geometry()

# Calculates area of holc graded tracts, m^2
tract_areas <- tracts %>%
  dplyr::filter(GEOID %in% holc$GEOID) %>%
  dplyr::mutate(tract_area = as.numeric(sf::st_area(geometry))) %>%
  dplyr::select(GEOID, tract_area) %>%
  st_drop_geometry()

# Calculates area of holc graded tracts for cities, m^2
holc_areas <- holc %>%
  dplyr::left_join(tract_areas, by = "GEOID") %>%
  dplyr::group_by(city, state) %>%
  dplyr::summarise(tract_area = sum(tract_area, na.rm = TRUE))
holc_areas <- as.data.frame(holc_areas)

# Table for cities which includes holc area, place area, and place to tract ratio
area_table <- holc_areas %>%
  dplyr::left_join(place_areas, by = c("city" = "CITY", "state" = "STUSPS.CITY")) %>%
  dplyr::mutate(place_tract_ratio = tract_area / place_area)
```

## Mapping and Statistics

### All Cities


Data for all cities indicates that while the new places data takes up the largest land area and includes a larger populace (roughly 50% more), the places where that data intersects with the new HOLC data has the densest population, likely making them more relevant for our analysis. This is not to discount them from importance however, as the places data is still more dense than the holc graded tracts, which may indicate a movement in population of certain areas.  


While outliers are removed because there are many of them (and they wreak havoc on the plot), the boxplots below indicate similar results for population density. The intersection of HOLC and places tracts have much higher density than either alone. When the intersection is included in both places and HOLC tracts, both come much closer to the intersection's density. We can conclude that going off of places is a good move, given that almost all HOLC tracts intersect with places tracts. With the move to places tracts, only 5% of land area and 2% of the population are discluded.


To further support this, 127 of the 202 cities have holc tracts which are completely enveloped in places.


```{r, echo = FALSE}
# Set general holc file to be holc
state_holc <- holc

# Dataframe of tracts with their geography and population; filtered to be only those in holc
state_holc_geo <- tracts %>%
  filter(GEOID %in% state_holc$GEOID) %>%
  select(GEOID, geometry) %>%
  distinct(GEOID, .keep_all = TRUE) %>%
  st_as_sf() %>%
  sf::st_set_crs(4269) %>%
  dplyr::left_join(all_acs_pops, by = "GEOID")

# Geography of combined place data for cities
state_place_geo <- holc_place %>%
  pull(geometry) %>%
  st_as_sf()

# Pulls list of places
pulled_places <- holc_geo %>%
  select(PLACE, STUSPS.PLACE) %>%
  distinct()

# Dataframe of distinct tracts in specified places with population estimate and geography
state_place_geo_tracts <- pulled_places %>%
  left_join(places_tracts, by = c("PLACE" = "NAME", "STUSPS.PLACE" = "STUSPS")) %>%
  select(GEOID, geometry) %>%
  distinct(GEOID, .keep_all = TRUE) %>%
  st_as_sf() %>%
  sf::st_set_crs(4269) %>%
  dplyr::left_join(all_acs_pops, by = "GEOID")

# Geography of intersecting tracts (places tracts and holc tracts) for city
holc_place_intersection <- state_holc_geo %>%
  filter(GEOID %in% state_place_geo_tracts$GEOID) %>%
    sf::st_set_crs(4269)

# Geography of combined places tracts without intersection
state_place_geo_tracts_new <- state_place_geo_tracts %>%
    dplyr::filter(!(GEOID %in% holc_place_intersection$GEOID))

# Geography of holc graded tracts without intersection
state_holc_geo_new <- state_holc_geo %>%
    dplyr::filter(!(GEOID %in% holc_place_intersection$GEOID))

# Number of cities which have holc graded tracts outside of combined place geography
state_holc_geo_aggregate <- holc %>%
  left_join(tracts %>% select(GEOID, geometry), by = "GEOID") %>%
  dplyr::filter(!(GEOID %in% holc_place_intersection$GEOID)) %>%
  dplyr::mutate(area = as.numeric(sf::st_area(geometry))) %>%
  dplyr::group_by(city, state) %>%
  dplyr::summarise(area = sum(area, na.rm = TRUE)) %>%
  dplyr::filter(area != 0) %>%
  dplyr::select(city, state) %>%
  nrow()

# Table of the total area, population, and population density of all groups
get_place_area_table <- function(spgtn, shgn, hpi, spgt, shg){
  # Combined area of tracts in each group
  tract_areas <- c(sum(as.numeric(sf::st_area(spgtn))),
         sum(as.numeric(sf::st_area(shgn))),
         sum(as.numeric(sf::st_area(hpi))),
         sum(as.numeric(sf::st_area(spgt))),
         sum(as.numeric(sf::st_area(shg))))
  
  # Proportion of total area of city which group takes up
  area_proportions <- c(tract_areas[1:3] / sum(tract_areas[1:3]) * 100, tract_areas[4:5] / sum(tract_areas[1:3]) * 100)
  
  # Combined population of tracts in each group
  tract_pops <- c(sum(spgtn$estimate, na.rm = TRUE),
         sum(shgn$estimate, na.rm = TRUE),
         sum(hpi$estimate, na.rm = TRUE),
         sum(spgt$estimate, na.rm = TRUE),
         sum(shg$estimate, na.rm = TRUE))
  
  # Proportion of total population of city which group takes up
  pop_proportions <- c(tract_pops[1:3] / sum(tract_pops[1:3]) * 100, tract_pops[4:5] / sum(tract_pops[1:3]) * 100)
  
  # Population density for each group by km^2, null values coerced to zero
  pop_density <- as.numeric(mapply("/", tract_pops, tract_areas, SIMPLIFY = FALSE))*1000000
  pop_density[is.na(pop_density)] <- 0
  
  # Names of columns
  df_cols <- c("Places Tracts (Exclusive)", "HOLC Tracts (Exclusive)", "Intersection", "Places Tracts (Inclusive)", 
               "HOLC Tracts (Inclusive)")
  
  # Creates dataframe where each row is a calculation and each column is a group. Rounded to 2 decimal places.
  place_area_table <- as.data.frame(do.call("rbind", list(paste0(round(area_proportions, 2), "%"),
                                                          paste0(round(pop_proportions, 2), "%"), 
                                                          round(pop_density, 2))),
                     row.names = c("Land Area", "Population", "Density (pop/km^2)"))
  colnames(place_area_table) <- df_cols
  return(place_area_table)
}

place_area_table <- get_place_area_table(spgtn = state_place_geo_tracts_new,
                                         shgn = state_holc_geo_new,
                                         hpi = holc_place_intersection, 
                                         spgt = state_place_geo_tracts, 
                                         shg = state_holc_geo)

# Creates a table visual
knitr::kable(place_area_table)

all_cities_plot <- ggplot() +
  geom_sf(data = state_place_geo_tracts_new, fill = "blue", alpha = 0.5) +
  geom_sf(data = state_holc_geo_new, fill = "green", alpha = 0.5) +
  geom_sf(data = holc_place_intersection, fill = "red", alpha = 0.5)

# Table of areas, populations, and densities for tracts in different groups
get_density_table <- function(spgtn, shgn, hpi, spgt, shg) {
  # Areas of tracts by group
  spgtn_area <- as.numeric(sf::st_area(spgtn))
  shgn_area <- as.numeric(sf::st_area(shgn))
  hpi_area <- as.numeric(sf::st_area(hpi))
  spgt_area <- as.numeric(sf::st_area(spgt))
  shg_area <- as.numeric(sf::st_area(shg))
  tract_areas <- c(spgtn_area, shgn_area, hpi_area, spgt_area, shg_area)
    
  # Types of tracts by group
  spgtn_type <- rep("Places (Exclusive)", length(spgtn_area))
  shgn_type <- rep("HOLC (Exclusive)", length(shgn_area))
  hpi_type <- rep("Intersection", length(hpi_area))
  spgt_type <- rep("Places (Inclusive)", length(spgt_area))
  shg_type <- rep("HOLC (Inclusive)", length(shg_area))
  tract_types <- c(spgtn_type, shgn_type, hpi_type, spgt_type, shg_type)
    
  # Population of tracts by group
  spgtn_pop <- spgtn$estimate
  shgn_pop <- shgn$estimate
  hpi_pop <- hpi$estimate
  spgt_pop <- spgt$estimate
  shg_pop <- shg$estimate
  tract_pops <- c(spgtn_pop, shgn_pop, hpi_pop, spgt_pop, shg_pop)
  
  # Density table with type, area m^2, pop, and density of tract km^2
  density_table <- as.data.frame(cbind(tract_types, tract_areas, tract_pops))
  colnames(density_table) <- c("type", "area", "pop")
  density_table <- density_table %>%
    dplyr::mutate(area = as.numeric(area),
                  pop = as.numeric(pop)) %>%
    replace(is.na(.), 0) %>%
    dplyr::mutate(density = pop / area * 1000000) %>%
    replace(is.na(.), 0)
  
  return(density_table)
}

density_table <- get_density_table(spgtn = state_place_geo_tracts_new,
                                   shgn = state_holc_geo_new,
                                   hpi = holc_place_intersection, 
                                   spgt = state_place_geo_tracts, 
                                   shg = state_holc_geo)

# Boxplots of density distributions for groups of tracts
all_cities_density_plot <- ggplot(density_table)+
  geom_boxplot(aes(y = density, x = type, fill = type), outlier.shape = NA)+
  stat_n_text(aes(y = density, x = type)) +
  scale_y_continuous(limits = quantile(density_table$density, c(0.1, 0.9)))

plot(all_cities_density_plot)

print(paste0("Cities where HOLC tracts are completely enshrouded in Place tracts: ", 202 - state_holc_geo_aggregate))

```

### Individual Cities

Investigating at the individual city level makes it possible to get an idea of not only how cities have grown over time, but also how redlining differs between geographies.

```{r, echo = FALSE}
# Table with 202 cities
city_table <- holc %>%
  dplyr::select(city, state) %>%
  dplyr::distinct() %>%
  dplyr::mutate(city_state = paste0(city, ", ", state))

# User input for city
inputPanel(
  shiny::selectInput("city", label = "City Name",
                     choices = city_table$city_state)
)

# Pull name of selected city
selected_city <- reactive({
  city_table %>%
    dplyr::filter(city_state == input$city) %>%
    dplyr::pull(city)
})

# Pull state of selected city
selected_state <- reactive({
  city_table %>%
    dplyr::filter(city_state == input$city) %>%
    dplyr::pull(state)
})

# Pull holc tracts for selected city
state_holc <- reactive({
  holc %>%
    dplyr::filter(city == selected_city(), state == selected_state())
})

# SF with tract geometry and population data for selected city
state_holc_geo <- reactive({
  tracts %>%
    dplyr::filter(GEOID %in% state_holc()$GEOID) %>%
    dplyr::select(GEOID, geometry) %>%
    dplyr::distinct(GEOID, .keep_all = TRUE) %>%
    dplyr::left_join(all_acs_pops, by = "GEOID") %>%
    sf::st_as_sf() %>%
    sf::st_set_crs(4269)
})

# SF with combined place geometry of city
state_place_geo <- reactive({
  holc_place %>%
    dplyr::filter(CITY == selected_city(), STUSPS.CITY == selected_state()) %>%
    dplyr::pull(geometry) %>%
    sf::st_as_sf()
})

# List of places in selected city
pulled_places <- reactive({
  holc_geo %>%
  dplyr::filter(CITY == selected_city(), STUSPS.CITY == selected_state()) %>%
  dplyr::select(PLACE, STUSPS.PLACE) %>%
  dplyr::distinct()
})

# SF of tracts with geography and populations in places
state_place_geo_tracts <- reactive({
  pulled_places() %>%
  dplyr::left_join(places_tracts, by = c("PLACE" = "NAME", "STUSPS.PLACE" = "STUSPS")) %>%
  dplyr::select(GEOID, geometry) %>%
  dplyr::distinct(GEOID, .keep_all = TRUE) %>%
  dplyr::left_join(all_acs_pops, by = "GEOID") %>%
  sf::st_as_sf() %>%
    sf::st_set_crs(4269)
})

# SF of tracts which are both holc graded and in place geographies
holc_place_intersection <- reactive({
  state_holc_geo() %>%
    dplyr::filter(GEOID %in% state_place_geo_tracts()$GEOID) %>%
    #dplyr::left_join(all_acs_pops, by = "GEOID") %>%
    sf::st_set_crs(4269)
})

# SF of place tracts without its intersection with holc tracts
state_place_geo_tracts_new <- reactive({
  state_place_geo_tracts() %>%
    dplyr::filter(!(GEOID %in% holc_place_intersection()$GEOID)) #%>%
    #dplyr::left_join(all_acs_pops, by = "GEOID")
})

# SF of holc tracts without its intersection with place tracts
state_holc_geo_new <- reactive({
  state_holc_geo() %>%
    dplyr::filter(!(GEOID %in% holc_place_intersection()$GEOID)) #%>%
    #dplyr::left_join(all_acs_pops, by = "GEOID")
})

# Table of the total area, population, and population density of all groups
place_area_table <- reactive({
  get_place_area_table(spgtn = state_place_geo_tracts_new(),
                       shgn = state_holc_geo_new(),
                       hpi = holc_place_intersection(), 
                       spgt = state_place_geo_tracts(), 
                       shg = state_holc_geo())
})

# Table of areas, populations, and densities for tracts in different groups
density_table <- reactive({
  get_density_table(spgtn = state_place_geo_tracts_new(),
                    shgn = state_holc_geo_new(),
                    hpi = holc_place_intersection(), 
                    spgt = state_place_geo_tracts(), 
                    shg = state_holc_geo())
})

```

```{r, echo=FALSE}

# Leaflet map for different groups in selected city. Includes popup with population.
renderLeaflet({
    leaflet() %>%
      addProviderTiles("CartoDB.Voyager") %>%
      addPolygons(data = state_holc_geo_new(),
                  weight = 0.5,
                  opacity = 0.5, 
                  color = "green",
                  popup = paste0("Tract: ", state_holc_geo_new()$GEOID, "<br>", "Population: ", state_holc_geo_new()$estimate)) %>%
      addPolygons(data = state_place_geo_tracts_new(),
                  weight = 0.5,
                  opacity = 0.5, 
                  color = "blue",
                  popup = paste0("Tract: ", state_place_geo_tracts_new()$GEOID, "<br>", "Population: ", state_place_geo_tracts_new()$estimate)) %>%
      addPolygons(data = holc_place_intersection(),
                  weight = 1,
                  opacity = 1,
                  color = "red",
                  popup = paste0("Tract: ", holc_place_intersection()$GEOID, "<br>", "Population: ", holc_place_intersection()$estimate)) %>%
      addLegend("bottomright", 
                colors = c("green", "blue", "red"), 
                labels = c("HOLC Tracts", "Place Tracts", "Intersection"),
                title = "Overlap of Tracts",
                opacity = 1)
  })

# Table of total area, population, and density of groups.
renderTable({
    place_area_table()
  },
  rownames = TRUE,
  colnames = TRUE)

# Boxplot of population density by groups
renderPlot({
  ggplot(density_table())+
  geom_boxplot(aes(y = density, x = type, fill = type))+
  stat_n_text(aes(y = density, x = type))
})

```

### Places with Pop > 50k

We also have additional data, in the form of individual places with a population of greater than 50k. This may be useful as well, but each state has different definitions of a place, documented [here](https://www2.census.gov/geo/pdfs/reference/GARM/Ch9GARM.pdf).

```{r, echo=FALSE}
# Pulls names of places with population greater than 50k
place_table <- places_pop_gt_50k %>%
  dplyr::select(PLACE, STUSPS, STATEFP.PLACE) %>%
  dplyr::distinct() %>%
  dplyr::mutate(place_state = paste0(PLACE, ", ", STUSPS))

# User input for place
inputPanel(
  shiny::selectInput("place", label = "Place Name",
                     choices = place_table$place_state)
)

# Pulls name of selected place
selected_place <- reactive({
  place_table %>%
    dplyr::filter(place_state == input$place) %>%
    dplyr::pull(PLACE)
})

# Pulls state of selected place
selected_place_state <- reactive({
  place_table %>%
    dplyr::filter(place_state == input$place) %>%
    dplyr::pull(STATEFP.PLACE)
})

# Pulls tracts for selected place with population
selected_place_geo <- reactive({
  places_tracts %>%
    dplyr::filter(NAME == selected_place() & STATEFP == selected_place_state()) %>%
    sf::st_as_sf() %>%
    sf::st_set_crs(4269) %>%
    dplyr::left_join(all_acs_pops, by = "GEOID")
})
```


```{r, echo=FALSE}
# Leaflet map for place. Includes popup with population.
renderLeaflet({
  leaflet() %>%
    addProviderTiles("CartoDB.Voyager") %>%
    addPolygons(data = selected_place_geo(),
                weight = 0.5,
                opacity = 0.5,
                color = "green",
                popup = paste0("Tract: ", selected_place_geo()$GEOID, "<br>", "Population: ", selected_place_geo()$estimate))
})
```
