---
title: "Trajectories"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
theme_set(theme_minimal())
options(scipen = 999)
```

## Objectives

-   Can we group / classify like cities within previously redlined areas?

```{r setup, include=FALSE}
redlining <-
  read.csv("raw data/All_Population_Weighted_HOLC_Grades_pt2_Threshold.csv")
redlining$GEOID <- str_pad(redlining$GEOID, width = 11, side = "left", pad = "0")
redlining$holc_grade_pop <- replace_na(redlining$holc_grade_pop, 0)

# Load LTDB Data
ltdb <- readRDS("outputs/ltdb_combined.RDS")

# Load ACS Data
acs_2012_processed <- readRDS("outputs/acs_2012.RDS") 

acs_2019_processed <- readRDS("outputs/acs_2019.RDS") 



ltdb <- bind_rows(ltdb, acs_2012_processed, acs_2019_processed)

ltdb$YEAR = as.numeric(ltdb$YEAR)



```

Idea:

-   Separate out the set of redlined cities in the LTDB
-   Examine trajectories in neighborhood indicators

Ex: How do the trajectories for Median Household Income compare between Redlined tracts and non-redlined tracts?

Following looks at Counties in Alabama, and changes in Median Household income for Black, White, and Total pops.

**Household Incomes**

```{r, message=FALSE}
household_income <- left_join(ltdb, redlining, by = "GEOID") %>%
  mutate(redlined = case_when(holc_grade_pop == 4 ~ "Redlined",
                              holc_grade_pop > 0 & holc_grade_pop < 4 ~ "Not Redlined",
                              TRUE ~ "Ungraded"),
         redlined_4 = case_when(holc_grade_pop >= 3.5 ~ "Best",
                                holc_grade_pop >= 2.5 & holc_grade_pop < 3.5 ~ "Still Desirable",
                                holc_grade_pop >= 1.5 & holc_grade_pop < 2.5 ~ "Definitely Declining",
                                holc_grade_pop < 1.5 & holc_grade_pop > 0 ~ "Hazardous", 
                                TRUE ~ "Ungraded")) %>%
  group_by(YEAR, STATE, COUNTY, redlined_4) %>%
  filter(redlined != "Ungraded") %>%
  summarize(Total = mean(`Median Household Income, Total`, na.rm = T),
            White = mean(`Median Household Income, White`, na.rm = T),
            Black = mean(`Median Household Income, Black`, na.rm = T)) %>%
  gather(key = "Median Household Income", value = "Dollars", c(Total, White, Black)) %>%
  ungroup()

household_income$redlined_4 <- factor(household_income$redlined_4, levels = c("Hazardous", "Definitely Declining", "Still Desirable", "Best", "Ungraded"))

household_income_plot <- function(State){
  household_income %>%
  filter(STATE == State) %>%
  ggplot(aes(x = YEAR, y = Dollars, color = `Median Household Income`))  +
  geom_point(aes(shape = `Median Household Income`)) +
  geom_line(aes(linetype = redlined)) +
  facet_wrap(~COUNTY)
}


```

```{r}
household_income_plot("Alabama")
household_income_plot("Massachusetts")

household_income_plot("Maryland")

household_income_plot("Wisconsin")

household_income %>%
  ggplot(aes(x = YEAR, y = Dollars, color = `Median Household Income`)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess") +
  scale_y_log10()+
  facet_grid(~redlined_4) +
  labs(y = "Dollars (Log Scale)")
```

**Education**

```{r, message=FALSE}
education <- left_join(ltdb, redlining, by = "GEOID") %>%
  mutate(redlined = case_when(holc_grade_pop == 4 ~ "Redlined",
                              holc_grade_pop > 0 & holc_grade_pop < 4 ~ "Not Redlined",
                              TRUE ~ "Ungraded")) %>%
  group_by(YEAR, STATE, COUNTY, redlined) %>%
  filter(redlined != "Ungraded") %>%
  summarize(`High School or Less` = mean(`Percent with High School Degree or Less`, na.rm = T),
            `College or More` = mean(`Percent with 4-year College Degree or More`, na.rm = T)) %>%
  gather(key = "Education", value = "Percent", c(`High School or Less`, `College or More`))  

education_plot <- function(State){
  education %>%
  filter(STATE == State) %>%
  ggplot(aes(x = YEAR, y = Percent, color = `Education`))  +
  geom_point(aes(shape = `Education`)) +
  geom_line(aes(linetype = redlined)) +
  facet_wrap(~COUNTY)
}

education_plot("Alabama")
education_plot("California")
education_plot("Massachusetts")
education_plot("Oregon")
education_plot("Colorado")
```

```{r}
education %>%
  ggplot(aes(x = YEAR, y = Percent, color = Education))  +
  geom_point( alpha = 0.4) +
  geom_smooth(method = "loess") +
  facet_grid(~redlined) +
  labs(title = "Trends in Education from 1970 to 2020")

```

## Poverty

```{r}
poverty <- left_join(ltdb, redlining, by = "GEOID") %>%
  mutate(redlined = case_when(holc_grade_pop == 4 ~ "Redlined",
                              holc_grade_pop > 0 & holc_grade_pop < 4 ~ "Not Redlined",
                              TRUE ~ "Ungraded")) %>%
  group_by(YEAR, STATE, COUNTY, redlined) %>%
  filter(redlined != "Ungraded") %>%
  summarize(Total = mean(`Percent in Poverty, Total`, na.rm = T),
            Black = mean(`Percent in Poverty, Black`, na.rm = T),
            White = mean(`Percent in Poverty, White`, na.rm = T),
            Native = mean(`Percent in Poverty, Native American`, na.rm = T),
            Hispanic = mean(`Percent in Poverty, Hispanic`, na.rm = T),
            `Asian/PI` = mean(`Percent in Poverty, Asian/PI`, na.rm = T)
            ) %>%
  gather(key = "Race / Ethnicity", 
         value = "Percent in Poverty", 
         c(Total, Black, White, Native, Hispanic, `Asian/PI`)) %>%
  mutate(`Race / Ethnicity` = factor(`Race / Ethnicity`, levels = c("Asian/PI", "Black", "Hispanic", "Native", "White", "Total")))

percent_in_poverty_plot <- function(State){
  poverty %>%
  filter(STATE == State,
         `Race / Ethnicity` %in% c("Black", "White", "Total")) %>%
  ggplot(aes(x = YEAR, y = `Percent in Poverty`, color = `Race / Ethnicity`))  +
  geom_point() +
  geom_line(aes(linetype = redlined)) +
  facet_grid( ~COUNTY)
}
```

```{r}
percent_in_poverty_plot("Alabama")
```

```{r}
poverty %>%
  filter(`Race / Ethnicity` %in% c("Black", "White", "Total")) %>%
  ggplot(aes(x = YEAR, y = `Percent in Poverty`, color = `Race / Ethnicity`)) + 
  geom_jitter(alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_grid(~ redlined) +
  labs(title = "Trends in Poverty from 1970 to 2020")
```

```{r}

```
