---
title: "Variance Exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(readxl)

theme_set(theme_minimal())
options(scipen = 999)

places_2020 <- read.csv("raw data/PLACES__Census_Tract_Data__GIS_Friendly_Format___2020_release.csv")

# 500 Cities data w/ Mental Health outcomes
cities_2019 <- read.csv("raw data/500_Cities__Local_Data_for_Better_Health__2019_release.csv")
cities_2018 <- read.csv("raw data/500_Cities__Local_Data_for_Better_Health__2018_release.csv")
cities_2017 <- read.csv("raw data/500_Cities__Local_Data_for_Better_Health__2017_release.csv")
cities_2017 <- cities_2017 %>%
  rename("PopulationCount" = "Population2010")

# Urban Displacement Project classifications of gentrification
udp <- read_xlsx("raw data/udp_data_spreadsheet_updated11062017.xlsx", sheet = 2)
```

# Objectives

-   Explore the confidence interval ranges for self-reported Poor Mental Health (as defined by "Mental health not good for \>=14 days among adults aged \>=18 Years")

-   Explore census tracts ranges throughout time

-   Use Urban Displacement Project (UDP) and Lisa Bates characterizations of gentrification for preliminary analysis

## Functions

```{r}
# Returns data set at tract level with mental health prevalence
process_500cities <- function(data){
  data %>%
  filter(!is.na(TractFIPS),
         Measure == "Mental health not good for >=14 days among adults aged >=18 Years") %>%
  mutate(conf_int_range = High_Confidence_Limit - Low_Confidence_Limit,
         PopulationCount_k = PopulationCount/1000)
}

# Plots Range in CI
plot_range_by_pop <- function(data){
  year <- parse_number(deparse(substitute(data)))
  
  data %>%
  process_500cities() %>%
  ggplot(aes(x = PopulationCount, y = conf_int_range, color = Data_Value)) +
  scale_color_binned(type = "viridis") +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "black", size = 0.6, se = T, na.rm = T, ) + 
  scale_x_log10() +
  ylim(0, 12.5) +
  labs(x = "Population Count (log scale)",
       y = "Confidence Interval Range",
       color = "Poor Mental Health Prevalence",
       title = paste("500 Cities Data -", year)) +
  theme(legend.position = "bottom")
}

# Plots range in Confidence intervals by Poor MH Prevalence
plot_range_by_prev <- function(data){
  year <- parse_number(deparse(substitute(data)))

  data %>%
    process_500cities() %>%
  ggplot(aes(y = conf_int_range, x = Data_Value)) +
  geom_point(aes(color = log(PopulationCount)),alpha = 0.2) +
  geom_smooth(method = "lm", color = "black", size = 0.6, se = T, na.rm = T, ) + 
  scale_color_binned(type = "viridis") +
  ylim(0,12.5) +
    xlim(0,40) +
  labs(size = "Population Count",
       y = "Confidence Interval Range",
       x = "Poor Mental Health Prevalence",
       title = paste("500 Cities Data -", year)) +

  theme(legend.position = "bottom")
}


```

## Single Year Plots of Variance Predictors

```{r}
plot_range_by_pop(cities_2017)
plot_range_by_pop(cities_2018)
plot_range_by_pop(cities_2019)
```

```{r}
plot_range_by_prev(cities_2017)
plot_range_by_prev(cities_2018)
plot_range_by_prev(cities_2019)

```

# Trend Analysis

```{r}
# Creating a smaller combined dataset of 500 Cities Data
subset_500 <- function(data){
  data %>%
    process_500cities() %>%
    select(Year, TractFIPS, StateAbbr, CityName, Data_Value, Low_Confidence_Limit, High_Confidence_Limit, PopulationCount)
}

subset_17 <- subset_500(cities_2017)
subset_18 <- subset_500(cities_2018)
subset_19 <- subset_500(cities_2019)

combined <- bind_rows(subset_17, subset_18, subset_19)

# Creating additional columns for Bates and UDP Typology, Range variable
combined <- combined %>%
  filter(TractFIPS %in% udp$FIPS) %>%
  mutate(TractFIPS = as.character(TractFIPS)) %>%
  left_join(udp, by = c("TractFIPS" = "FIPS")) %>%
  separate(UDP_TYPE.15,
           into = c("Income Group 2015", "Status 2015"),
           sep = "_") %>%
  mutate(
    `Income Group 2015` = ifelse(`Income Group 2015` == "LI", "Low Income", "Medium or High"),
    `Status 2015` = case_when(
      `Status 2015` == "AG" ~ "Advanced Gentrification",
      `Status 2015` == "AtRisk" ~ "At Risk",
      `Status 2015` == "NL" ~ "Not Losing",
      `Status 2015` == "UG" ~ "Undergoing Gentrification",
      TRUE ~ ""
    ),
    `Status 2015` = ifelse(`Income Group 2015` == "Medium or High", NA, `Status 2015`),
    `Status 2015` = factor(
      `Status 2015`,
      levels = c(
        "Not Losing",
        "At Risk",
        "Undergoing Gentrification",
        "Advanced Gentrification"
      )
    ),
    Range = High_Confidence_Limit - Low_Confidence_Limit,
    NH_Typo15 = factor(
      NH_Typo15,
      levels = c(
        "Susceptible",
        "Early: Type1",
        "Early: Type2",
        "Dynamic",
        "Late",
        "Continued Loss"
      )
    )
  )

```

## UDP Gentrification Determination

Gentrification applicable only to subset of low-income tracts

| Low             | Medium or High  |
|-----------------|-----------------|
| Not Losing to G | Not Losing to G |
| At Risk G       | At Risk G       |
| Undergoing G    | Undergoing G    |
| Advanced G      | Advanced G      |

```{r}
# Plot of Trend
combined %>%
  ggplot(aes(x = factor(Year), y = Data_Value, group = TractFIPS, color = `Status 2015`)) +
  geom_point(aes(size = Range), alpha = 0.4) +
  geom_line() +
  labs(subtitle = "UDP Gentrification Typology",
       x = "Year", 
       y = "Poor Mental Health Prevalence")  +
  facet_grid(~`Status 2015`, labeller = label_wrap_gen()) + 
  theme(legend.position = "bottom",
        legend.box = "vertical") +
  guides(color = FALSE)

# Plot of Variance
combined %>%
  ggplot(aes(x = factor(Year), y = Range, group = TractFIPS, color = `Status 2015`)) +
  geom_point(aes(size = Data_Value), alpha = 0.4) +
  geom_line() +
  labs(subtitle = "UDP Gentrification Typology",
       x = "Year", 
       y = "Range",
       size = "Poor Mental Health Prevalence") +
  facet_grid(~`Status 2015`) +
  theme(legend.position = "bottom",
        legend.box = "vertical") +
  guides(color = FALSE)
```

## Lisa Bates Gentrification Typology

| Stage          | Vulnerable Population | Demographic Change | Housing Market Condition |
|----------------|-----------------------|--------------------|--------------------------|
| Susceptible    |                       |                    |                          |
| Early Type 1   |                       |                    |                          |
| Early Type 2   |                       |                    |                          |
| Dynamic        |                       |                    |                          |
| Late           |                       |                    |                          |
| Continued Loss |                       |                    |                          |

```{r}
# Plot of Trend
combined %>%
  ggplot(aes(x = factor(Year), y = Data_Value, group = TractFIPS, color = NH_Typo15)) +
  geom_point(aes(size = Range), alpha = 0.4) +
  geom_line() +
  labs(x = "Year", y = "Poor Mental Health Prevalence",
       subtitle = "Lisa Bates Gentrification Typology")  +
  facet_grid(~NH_Typo15) + 
  theme(legend.position = "bottom",
        legend.box = "vertical") +
  guides(color = FALSE)

# Plot of Variance
combined %>%
  ggplot(aes(x = factor(Year), y = Range, group = TractFIPS, color = NH_Typo15)) +
  geom_point(aes(size = Data_Value), alpha = 0.4) +
  geom_line() +
  labs(x = "Year", y = "Range",
       subtitle = "Lisa Bates Gentrification Typology",
       size = "Poor Mental Health Prevalence")  +
  facet_grid(~NH_Typo15) + 
  theme(legend.position = "bottom",
        legend.box = "vertical") +
  guides(color = FALSE)
```

## Miscellaneous

```{r}
combined %>%
  filter(TractFIPS %in% udp$FIPS) %>%
  mutate(TractFIPS = as.character(TractFIPS)) %>%
  left_join(udp, by = c("TractFIPS" = "FIPS")) %>%
  separate(UDP_TYPE.15, into = c("Income Group 2015", "Status 2015"), sep = "_")%>%
  mutate(`Income Group 2015` = ifelse(`Income Group 2015` == "LI", "Low Income", "Medium or High"),
         `Status 2015` = case_when(`Status 2015` == "AG" ~ "Advanced Gentrification",
                                   `Status 2015` == "AtRisk" ~ "At Risk",
                                   `Status 2015` == "NL" ~ "Not Losing",
                                   `Status 2015` == "UG" ~ "Undergoing Gentrification",
                                   TRUE ~ ""),
         `Status 2015` = factor(`Status 2015`, levels = c("Not Losing", "At Risk", "Undergoing Gentrification", "Advanced Gentrification")),
         Range = High_Confidence_Limit - Low_Confidence_Limit) %>%
  filter(`Income Group 2015` == "Low Income") %>%
  ggplot(aes(x = Range)) +
  geom_histogram(aes(fill = `Status 2015`), alpha = 0.5, color = "white") +
  facet_grid(rows = vars(`Status 2015`), labeller = label_wrap_gen()) +
  theme(panel.grid.minor = element_blank(),
        legend.position = "none") 
  
```

```{r}

```

```{r}
dat <- combined %>%
  filter(TractFIPS %in% udp$FIPS) %>%
  mutate(TractFIPS = as.character(TractFIPS)) %>%
  left_join(udp, by = c("TractFIPS" = "FIPS")) %>%
  separate(UDP_TYPE.15, into = c("Income Group 2015", "Status 2015"), sep = "_")%>%
  mutate(`Income Group 2015` = ifelse(`Income Group 2015` == "LI", "Low Income", "Medium or High"),
         `Status 2015` = case_when(`Status 2015` == "AG" ~ "Advanced Gentrification",
                                   `Status 2015` == "AtRisk" ~ "At Risk",
                                   `Status 2015` == "NL" ~ "Not Losing",
                                   `Status 2015` == "UG" ~ "Undergoing Gentrification",
                                   TRUE ~ ""),
         `Status 2015` = factor(`Status 2015`, levels = c("Not Losing", "At Risk", "Undergoing Gentrification", "Advanced Gentrification")),
         Range = High_Confidence_Limit - Low_Confidence_Limit) %>%
  filter(`Income Group 2015` == "Low Income") 

mod <- lm(Range ~ Data_Value + NH_Typo15 + PopulationCount, data = dat)
summary(mod)
```
